Tools and Ideas for the Proceedural Generation of Shadow Sculptures.

### 2015-12-15 paper/paper.tex
Finished draft of appendix A.
Covers everything that should be in the appendix. Variable naming in this derivation is not consistent with the variable naming in rayTracing.py.

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[toc,page]{appendix}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

@@

The simplest method of representing nodes and edges would be to keep the number of both constant, moving the nodes around while retaining the same edge mappings for all genomes.  This way, one could represent each node as an unsigned integer, or an array of three unsigned integers representing the positive incremental distance from some origin point.  Given the nature of the problem, one can disregard any space not within the projection from the light source to a given section of the wall.  This allows the bounding of the entire system in a cube, which we can use to define our \texttt{0} and our \texttt{MAX\_INT} for each coordinate direction.  Next, we would create an array of unsigned integers, of length 3 times the number of nodes in the system.  This would be the genome for the genetic algorithm.  The fitness function would be defined as the number of matching pixels (within a certain threshold) between the input and the generated shadow using one of the methods above.  Using such a method would prevent situations where the graph became unconnected, but it does not address issues of structural integrity any further than this.  Since the nature of the problem allows for many local solutions, the ability of such a genetic algorithm to converge may be slowed.

@@

@@

The first step is to derive the equation of the shadow cone. A generic point on the cone will be denoted by $\vec{x}$. The axis of the cone passes through $\vec{l}$ and through $\vec{c}$. The defining property of $\vec{x}$ is 
\begin{equation}\label{xproperty}
\|orth_{\vec{c}-\vec{l}} (\vec{x} - \vec{l})\| = \|proj_{\vec{c}-\vec{l}} (\vec{x} - \vec{l})\| \cdot k
\end{equation}
where $k$ is a constant characteristic of the aperture of the cone. Equation (\ref{xproperty}) says that the distance of a point from the cone's axis is in constant proportion to its distance along the axis. The value of $k$ can be calculated from any point on the cone. We will calculate $k$ from the projection and orthogonal projection of a point of tangency, $\vec{p}$, of the sphere and cone.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{sphere_shadow_figure_1.png}
    \caption{Depicted is the light source, a sphere, and a ray of light tangent to the sphere (labeled as generatrix of cone).}
    \label{fig:shadow_1}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{sphere_shadow_figure_2.png}
    \caption{This figure includes in addition the orthogonal projection from the point of tangency to the axis of the cone. The distance $L''$ is from the light source to the tip of the projection vector.}
    \label{fig:shadow_2}
\end{figure}

@@

Now we use the defintions of $orth_{\vec{u}} \vec{v}$ and $proj_{\vec{u}} \vec{v}$ to simplify equation (\ref{xproperty}).
\begin{align*}
proj_{\vec{u}} \vec{v} = \vec{u} \frac{\vec{u}\cdot\vec{v}}{\vec{u}\cdot\vec{u}}\\
orth_{\vec{u}} \vec{v} = \vec{v} - proj_{\vec{u}} \vec{v}
\end{align*}
Start by squaring both sides of \ref{xproperty} and substituting. Let $\vec{d} = \vec{c}-\vec{l}$, $y = \vec{x}-\vec{l}$.
\begin{align*}
\|\vec{y} - \vec{d} \frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}}\|^2 = \|\vec{d} \frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}}\|^2 \cdot k^2\\
\vec{y}\cdot\vec{y}-2\vec{y}\cdot\vec{d} \frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}}+\vec{d}\cdot\vec{d}(\frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}})^2 = (\vec{d}\cdot\vec{d}) (\frac{\vec{d}\cdot\vec{y}}{\vec{d}\cdot\vec{d}})^2 \cdot k^2
\end{align*}
Multiply each side by $\vec{d}\cdot\vec{d}$:
\begin{align*}
(\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})-2(\vec{d}\cdot\vec{y})^2+(\vec{d}\cdot\vec{y})^2  = (\vec{d}\cdot\vec{y})^2 \cdot k^2\\
(\vec{y}\cdot\vec{y})(\vec{d}\cdot\vec{d})-(\vec{d}\cdot\vec{y})^2  = (\vec{d}\cdot\vec{y})^2 \cdot k^2
\end{align*}
Now we find the intersection of this cone with the $z=0$ plane.

>>>>

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[toc,page]{appendix}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{breqn}

@@

The simplest method of representing nodes and edges would be to keep the number of both constant, moving the nodes around while retaining the same edge mappings for all genomes.  This way, one could represent each node as an unsigned integer, or an array of three unsigned integers representing the positive incremental distance from some origin point.  Given the nature of the problem, one can disregard any space not within the projection from the light source to a given section of the wall.  This allows the bounding of the entire system in a cube, which we can use to define our 0 and our \texttt{MAX\_INT} for each coordinate direction.  Next, we would create an array of unsigned integers, of length 3 times the number of nodes in the system.  This would be the genome for the genetic algorithm.  The fitness function would be defined as the number of matching pixels (within a certain threshold) between the input and the generated shadow using one of the methods above.  Using such a method would prevent situations where the graph became unconnected, but it does not address issues of structural integrity any further than this.  Since the nature of the problem allows for many local solutions, the ability of such a genetic algorithm to converge may be slowed.

@@

We should use bresenham's algorithm to draw an ellipse once we have its parameters, instead of the naive method we use now.

@@

The first step is to derive the equation of the shadow cone. A generic point on the cone will be denoted by $\vec{s}$. The axis of the cone passes through $\vec{l}$ and through $\vec{c}$. The defining property of $\vec{s}$ is 
\begin{equation}\label{sproperty}
\|orth_{\vec{c}-\vec{l}} (\vec{s} - \vec{l})\| = \|proj_{\vec{c}-\vec{l}} (\vec{s} - \vec{l})\| \cdot k
\end{equation}
where $k$ is a constant characteristic of the aperture of the cone. Equation (\ref{sproperty}) says that the distance of a point from the cone's axis is in constant proportion to its distance along the axis. The value of $k$ can be calculated from any point on the cone. We will calculate $k$ from the projection and orthogonal projection of a point of tangency, $\vec{p}$, of the sphere and cone.

@@

\begin{figure}[h]\label{fig:shadow_1}
    \centering
    \includegraphics[width=0.8\textwidth]{sphere_shadow_figure_1.png}
    \caption{Depicted is the light source, a sphere, and a ray of light tangent to the sphere (labeled as generatrix of cone).}
\end{figure}

\begin{figure}[h]\label{fig:shadow_2}
    \centering
    \includegraphics[width=0.8\textwidth]{sphere_shadow_figure_2.png}
    \caption{This figure includes in addition the orthogonal projection from the point of tangency to the axis of the cone. The distance $L''$ is from the light source to the tip of the projection vector.}
\end{figure}

@@

Now we use the defintions of $orth_{\vec{u}} \vec{v}$ and $proj_{\vec{u}} \vec{v}$ to simplify equation (\ref{xproperty}).
\begin{align*}
proj_{\vec{u}} \vec{v} &= \vec{u} \frac{\vec{u}\cdot\vec{v}}{\vec{u}\cdot\vec{u}}\\
orth_{\vec{u}} \vec{v} &= \vec{v} - proj_{\vec{u}} \vec{v}
\end{align*}
Start by squaring both sides of \ref{sproperty} and substituting. Let $\vec{d} = \vec{c}-\vec{l}$, $\vec{t} = \vec{s}-\vec{l}$.
\begin{align*}
\|\vec{t} - \vec{d} \frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}}\|^2 &= \|\vec{d} \frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}}\|^2 \cdot k^2\\
\vec{t}\cdot\vec{t}-2\vec{t}\cdot\vec{d} \frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}}+\vec{d}\cdot\vec{d}(\frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}})^2 &= (\vec{d}\cdot\vec{d}) (\frac{\vec{d}\cdot\vec{t}}{\vec{d}\cdot\vec{d}})^2 \cdot k^2
\end{align*}
Multiply each side by $\vec{d}\cdot\vec{d}$:
\begin{align*}
(\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})-2(\vec{d}\cdot\vec{t})^2+(\vec{d}\cdot\vec{t})^2  &= (\vec{d}\cdot\vec{t})^2 \cdot k^2\\
(\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})-(\vec{d}\cdot\vec{t})^2 & = (\vec{d}\cdot\vec{t})^2 \cdot k^2
\end{align*}
Divide both sides by $k^2$ and substitute $k^2 = r^2 / ( L^2-r^2)$.
\begin{align*}
((\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})-(\vec{d}\cdot\vec{t})^2)\cdot ( L^2-r^2)/r^2 &=  (\vec{d}\cdot\vec{t})^2\\
(\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})( L^2-r^2)/r^2 &=  (\vec{d}\cdot\vec{t})^2(1 + ( L^2-r^2)/r^2)\\
(\vec{t}\cdot\vec{t})(\vec{d}\cdot\vec{d})( L^2-r^2)/r^2 &=  (\vec{d}\cdot\vec{t})^2 L^2/r^2
\end{align*}
The quantities $\vec{d}\cdot\vec{d}$ and $L^2$ are equal (see figure \ref{fig:shadow_1}).
\begin{equation}\label{coneeqn}
(\vec{t}\cdot\vec{t})( L^2-r^2) =  (\vec{d}\cdot\vec{t})^2
\end{equation}
Now we find the intersection of this cone with the $z=0$ plane. Let $x$ and $y$ be coordinates on the shadow plane. The condition for intersection is that the third component of $\vec{s}$ be $0$. We put $\vec{t}$ and $\vec{d}$ in terms of $x$ and $y$ and substitute into equation \ref{coneeqn}.
\begin{align*}
\vec{s}& = \langle x,y,0 \rangle\\
\vec{d}\cdot\vec{t} &= \vec{d}\cdot\vec{s} - \vec{d}\cdot\vec{l} = d_1x+d_2y - \vec{d}\cdot\vec{l}\\
\vec{t}\cdot\vec{t} &= \vec{s}\cdot\vec{s} -2\vec{s}\cdot\vec{l} + \vec{l}\cdot\vec{l} = x^2 + y^2-2l_1x-2l_2y+ \vec{l}\cdot\vec{l}\\
(\vec{d}\cdot\vec{t})^2 &= (d_1x+d_2y - \vec{d}\cdot\vec{l})^2\\
&= (d_1x+d_2y)^2 - 2(d_1x+d_2y)(\vec{d}\cdot\vec{l}) + (\vec{d}\cdot\vec{l})^2\\
&= d_1^2x^2+2d_1d_2xy+d_2^2y^2-2d_1x(\vec{d}\cdot\vec{l})-2d_2y(\vec{d}\cdot\vec{l})+(\vec{d}\cdot\vec{l})^2\\
&= d_2^2y^2 + (2d_1d_2x -2d_2(\vec{d}\cdot\vec{l}))y + (d_1^2x^2-2d_1x(\vec{d}\cdot\vec{l})+(\vec{d}\cdot\vec{l})^2)\\
(\vec{t}\cdot\vec{t})( L^2-r^2) &= (x^2 + y^2-2l_1x-2l_2y+ \vec{l}\cdot\vec{l})(L^2-r^2)\\
&= (L^2-r^2)y^2 + (-2l_2(L^2-r^2))y + (x^2-2l_1x+ \vec{l}\cdot\vec{l})(L^2-r^2)
\end{align*}
For the final result we subtract the one side of equation \ref{coneeqn} from the other to obtain:
\begin{dmath*}
0 = (\vec{d}\cdot\vec{t})^2 - (\vec{t}\cdot\vec{t})( L^2-r^2)\\
= (d_2^2 - (L^2-r^2))y^2 + (2d_1d_2x -2d_2(\vec{d}\cdot\vec{l})+2l_2(L^2-r^2))y+(d_1^2x^2-2d_1x(\vec{d}\cdot\vec{l})+(\vec{d}\cdot\vec{l})^2 - (x^2-2l_1x+ \vec{l}\cdot\vec{l})(L^2-r^2))
\end{dmath*}
